---
title: "Data Description"
author: "Billy Johnson"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load in the packages
```{r}
library(neonUtilities)
library(tidyverse)
library(dplyr)
library(kableExtra)
```

# Load in the Data for Macroinvertebrates
```{r}
# Invertebrate Data
load("~/Desktop/Colorado State University Class Things/SUPER PROJECT/SUPER Project/DataMacro/invert.RData")

# Clean Invertebrate Data
load("~/Desktop/Colorado State University Class Things/SUPER PROJECT/SUPER Project/DataMacro/CleanInvert.RData")
```

1. inv_persample: Aquatic macroinvertebrate habitat data and QC metrics per sample
2. inv_identificationHistory: Aquatic macroinvertebrate identification history for records where identifications
have changed
3. inv_fieldData: Aquatic macroinvertebrate field data
4. inv_pervial: Aquatic macroinvertebrate identified archive data
5. inv_taxonomyProcessed: Aquatic macroinvertebrate identifications by expert taxonomists - desynonymized
6. inv_taxonomyRaw: Aquatic macroinvertebrate identifications by expert taxonomists - raw
7. variables: Description and units for each column of data in data tables
8. readme: Data product description, issue log, and other metadata about the data product
9. validation: Description of data validation applied at the points of collection and ingest


# Create tibbles from the list for field data and taxonomy data
```{r}
invert.tax <-
  invert[[5]] %>% 
  as_tibble()

str(invert.tax)


invert.field <-
  invert[[3]] %>% 
  as_tibble()

str(invert.field)
```

Macroinvertebrate Data
```{r}
invert.all <-
  inner_join(invert.tax, invert.field, by = "sampleID")

str(invert.all)
```

```{r}
NEONFILE.CLEAN2 <- invert.all %>% 
  select(collectDate.x,sampleID, scientificName, estimatedTotalCount, benthicArea, siteID.x) %>% 
  separate(collectDate.x, c("collectDate", "time"), sep = " ") %>% 
  mutate(Sample = substr(sampleID, 1, 11)) %>% 
  group_by(collectDate, time, scientificName, siteID.x ) %>% 
  filter(scientificName %in% c("Plecoptera sp.", "Trichoptera sp.", ("Ephemeroptera sp."))) %>% 
  summarize(inv.density = sum(estimatedTotalCount)/sum(benthicArea)) 

kable(NEONFILE.CLEAN2)
save(NEONFILE.CLEAN2, file = "dataMacro/CleanInvert.RData")

```

# Show the distribution of species density collected
```{r}
NEONFILE.CLEAN2 %>% 
  ggplot(aes(x = collectDate.x, y = inv.density, color = scientificName))+
  geom_point()
```

# Box plot of species 
```{r}
NEONFILE.CLEAN2 %>% 
  ggplot(aes(x = scientificName, y = inv.density, color = scientificName))+
  geom_boxplot()
```


# Load in the data for the Water Quality
```{r}
waterqual <- read_csv("DataWaterQuality/waterqual.csv")

```

# Clean the water quality data
```{r}
clean_wq <- waterqual %>% 
  rename(collectDate.x = startDateTime)
```

# Combine the macroinvertebrate data and the water quality data
```{r}
combined.data <-
  left_join(NEONFILE.CLEAN2, waterqual, by = )
```

