---
title: "SUPER_nutrient_data"
author: "Gabriella Vieira"
date: "2023-11-16"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# SECTION 1: DATA CLEANING AND ORGANIZATION 

## Load and read data


### Packages

```{r}
install.packages('dataRetrieval')
install.packages("tidyverse")
install.packages("flextable")
install.packages("ggdark")
install.packages("ggtheme")
install.packages(c("httr", "jsonlite"))
install.packages("neonUtilities")
```


```{r}
library(tibble)
library(rlang)
library(tidyverse) #which includes ggplot2
library(tmap) # interactive and static maps
library(sf) # to manage spatial data to make maps
library(tigris) # to import census tract spatial data
library(flextable)
library(ggdark)
library(ggplot2)
library(viridis)
library(RColorBrewer)
library(ggthemes)
library(terra)
library(mapproj)
library(dataRetrieval)
library(httr)
library(jsonlite)
library(neonUtilities)
```

Read data

```{r}
monthprecip_raw <- read.csv("data//PRISM_precip.csv")
watertemp_clean <- read.csv("data//watertemp_clean.csv")
```



#STREAM DISCHARGE DATA

Parameter metadata
```{r}
parameterCd <- "00060"
parameterINFO <- readNWISpCode(parameterCd)
```

```{r}
dis_attr <- attr(discharge, "variableInfo")
```

09032990
, "09026500", "06696980"

#USGS discharge
```{r}
siteNumber <- "09022000"
parameterCd <- "00060" # Discharge
startDate <- "2008-11-01"
endDate <- "2023-11-01"

discharge_usgs <- readNWISdv(siteNumber, parameterCd, startDate, endDate)
```


#NEON discharge
```{r}
discharge <- loadByProduct(dpID = "DP4.00130.001",
                              site = c("ARIK", "COMO", "WLOU"),
                              startdate = "2016-01",
                              enddate = "2023-01",
                              check.size = F)
```


Tibble the water temperature data
```{r}
dischargeNeon <- as_tibble(discharge$csd_continuousDischarge[, c("siteID", "endDate", "equivalentStage", "nonSystematicUnc", "systematicUnc", "stageUnc","maxpostDischarge", "withParaUncQLower2Std", "withParaUncQUpper2Std", "withRemnUncQLower2Std", "withRemnUncQUpper2Std", "dischargeFinalQF", "dischargeFinalQFSciRvw")])
```

Exporting to CSV file
```{r}
write.csv(dischargeNeon, "data//dischargeNeon.csv", row.names = FALSE)
```


Box plot of mean water temp per site per year
```{r}
ggplot(watertemp_clean, aes(x = Year, y = surfWaterTempMean)) +
  geom_boxplot(fill = "blue", alpha = 0.7) +
  facet_wrap(~siteID, scales = "free_y", ncol = 1) +
  labs(title = "Boxplots of surfWaterTempMean by Year and siteID",
       x = "Year",
       y = "surfWaterTempMean") +
  theme_minimal()
```


#PRECIPITATION AND MEAN TEMPERATURE DATA


graph precipitation data (filtered by date just to see better)
```{r}
monthprecip_raw %>% 
  filter(Date >= "2021-01") %>%
  ggplot(aes(x = Date, y = ppt..mm.)) +
  geom_point(color = "black")
```


graph temperature data
```{r}
monthprecip_raw %>% 
  filter(Date >= "2020-01") %>%
  ggplot(aes(x = Date, y = tmean..degrees.C.)) +
  geom_point(color = "black")
```



Surface water temperature
```{r}
Watertemp <- loadByProduct(dpID = "DP1.20053.001",
                              site = c("ARIK", "COMO", "WLOU"),
                              startdate = "2016-01",
                              enddate = "2023-01",
                              check.size = F)
```


Tibble the water temperature data
```{r}
water_temp30 <- as_tibble(Watertemp$TSW_30min[, c("siteID", "startDateTime", "endDateTime", "surfWaterTempMean", "surfWaterTempMinimum", "surfWaterTempMaximum", "surfWaterTempVariance")])
```

```{r}
watertemp_clean <- water_temp30[complete.cases(water_temp30$surfWaterTempMean), ]

```

Exporting to CSV file
```{r}
write.csv(watertemp_clean, "data//watertemp_clean.csv", row.names = FALSE)
```



#NUTRIENT DATA


All nutrients
```{r}
Waterquality <- loadByProduct(dpID = "DP1.20288.001",
                              site = c("ARIK", "COMO", "WLOU"),
                              startdate = "2016-01",
                              enddate = "2023-01",
                              check.size = F)
```

Tibble the water quality data
```{r}
water_qual <- as_tibble(Waterquality$waq_instantaneous[, c("siteID", "startDateTime", "endDateTime", "specificConductance","specificCondFinalQF", "dissolvedOxygen", "pH", "chlorophyll", "turbidity", "fDOM")])
```

```{r}
waterqual_clean <- water_qual[complete.cases(water_qual$specificConductance), ]
```

Exporting to CSV file
```{r}
write.csv(waterqual_clean, "data//waterqual.csv", row.names = FALSE)
```


```{r}
write.csv(Waterquality$waq_instantaneous, "data//waterqual_raw.csv", row.names = FALSE)
```





#Nitrate data

```{r}
nitrate <- loadByProduct(dpID = "DP1.20033.001",
                              site = c("ARIK", "COMO", "WLOU"),
                              startdate = "2016-01",
                              enddate = "2023-01",
                              check.size = F)
```


Tibble nitrate data
```{r}
nitrate_raw <- as_tibble(nitrate$NSW_15_minute[, c("siteID", "startDateTime", "endDateTime", "surfWaterNitrateMean","surfWaterNitrateMinimum", "surfWaterNitrateMaximum", "surfWaterNitrateVariance")])
```

```{r}
nitrate_clean <- nitrate_raw[complete.cases(nitrate_raw$surfWaterNitrateMean), ]
```

Exporting to CSV file
```{r}
write.csv(nitrate_clean, "data//nitrate.csv", row.names = FALSE)
```


